---
title: "Untitled"
author: "Chase Clark"
date: "January 7, 2018"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(FactoMineR)
library(factoextra)
library(rgl)
library(plotly)
library(Rtsne)
library(dendextend)
library(networkD3)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

proteinMatrix<-readRDS("data/ProteinMatrixWithDefaultIDBacSettings.rds")

```



# PCA Analysis


We need to find all the samples for which there is 16s-assigned taxonomic names:

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sampleNames <- row.names(proteinMatrix)

spl<-strsplit(sampleNames,"-")
lspl<-sapply(spl,length)
# Find which have taxonomic names
three<-which(lspl==3)
taxonomicNames<-sapply(spl[three],function(x)x[[3]])
taxonomicNames

```


Just return Genus

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
genera<-unlist(lapply(taxonomicNames,function(x)strsplit(x," ")[[1]][[1]]))
genera
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
length(unique(genera))
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
tibs<-as_tibble(genera)
top5<- tibs %>% group_by(value) %>% tally(sort=TRUE)
# display table of genera sorted by prevalence
 tibs %>% group_by(value) %>% tally(sort=TRUE)

```



Now we need to make a column for colors
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}


cols2<-rep("black",length(sampleNames))
cols2[grep(top5$value[[1]],sampleNames)]<-"red"
cols2[grep(top5$value[[2]],sampleNames)]<-"green"
cols2[grep(top5$value[[3]],sampleNames)]<-"blue"
cols2[grep(top5$value[[4]],sampleNames)]<-"orange"
cols2[grep(top5$value[[5]],sampleNames)]<-"purple"





cols<-rep("Other",length(sampleNames))
cols[grep(top5$value[[1]],sampleNames)]<-top5$value[[1]]
cols[grep(top5$value[[2]],sampleNames)]<-top5$value[[2]]
cols[grep(top5$value[[3]],sampleNames)]<-top5$value[[3]]
cols[grep(top5$value[[4]],sampleNames)]<-top5$value[[4]]
cols[grep(top5$value[[5]],sampleNames)]<-top5$value[[5]]


```






PCA with scaling, but no log normaliztion
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

pcaResults<- PCA(proteinMatrix,graph=FALSE,scale.unit = T)
# Get the first 5 dimensions
fiveDimensions <- as.data.frame(pcaResults$ind$coord)

forPlotting <- as_tibble(cbind(fiveDimensions,sampleNames))

p <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_text()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- No Scaling, No log normalization")+
    theme(plot.title=element_text(size=15),legend.position="none")

p2 <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_point()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- Scaled, No log normalization")



ggplotly(p)
ggplotly(p2)


```




PCA without scaling, no log normaliztion



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
pcaResults<- PCA(proteinMatrix,graph=FALSE,scale.unit = F)
# Get the first 5 dimensions
fiveDimensions <- as.data.frame(pcaResults$ind$coord)
sampleNames <- row.names(fiveDimensions)
forPlotting <- as_tibble(cbind(fiveDimensions,sampleNames))

p <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_text()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- No Scaling, No log normalization")+
    theme(plot.title=element_text(size=15),legend.position="none")

p2 <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_point()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- No Scaling, No log normalization")

ggplotly(p)
ggplotly(p2)

```






PCA with scaling,  with log normaliztion


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}


sampleNames <- row.names(proteinMatrix)

zeroHandle <-sapply(as_tibble(proteinMatrix),function(x){
          unlist(lapply(x,function(y){
              if (y<0.1){     
              
              y<- .001   
              } else{
                  y
              }
          }
          ))
      }
  )



proteinMatrixTibble<-as_tibble(zeroHandle)

proteinMatrixTibble <- proteinMatrixTibble %>% mutate_all(log10)

pcaResults<- PCA(proteinMatrixTibble,graph=FALSE,scale.unit = T)
# Get the first 5 dimensions
fiveDimensions <- as.data.frame(pcaResults$ind$coord)

forPlotting <- as_tibble(cbind(fiveDimensions,sampleNames))


p <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_text()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- Scaled and log normalized")+
    theme(plot.title=element_text(size=15),legend.position="none")

p2 <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_point()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- Scaled and log normalized")


ggplotly(p)
ggplotly(p2)




```






PCA without scaling,  with log normalization



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

sampleNames <- row.names(proteinMatrix)

zeroHandle <-sapply(as_tibble(proteinMatrix),function(x){
          unlist(lapply(x,function(y){
              if (y<0.1){     
              
              y<- .001   
              } else{
                  y
              }
          }
          ))
      }
  )



proteinMatrixTibble<-as_tibble(zeroHandle)

proteinMatrixTibble <- proteinMatrixTibble %>% mutate_all(log10)

pcaResults<- PCA(proteinMatrixTibble,graph=FALSE,scale.unit = F)
# Get the first 5 dimensions
fiveDimensions <- as.data.frame(pcaResults$ind$coord)

forPlotting <- as_tibble(cbind(fiveDimensions,sampleNames))


p <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_text()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- No scaling, Log normalized")+
    theme(plot.title=element_text(size=15),legend.position="none")

p2 <- ggplot(forPlotting,aes(Dim.1,Dim.2,label=sampleNames,color=cols))+
    geom_point()+
    xlab("Dimension 1")+
    ylab("Dimension 2")+
    ggtitle(" PCA, Protein Data- No scaling, Log normalized")



ggplotly(p)
ggplotly(p2)



```

### Look at PCA results in 3D- PCA without scaling,  with log normaliztion

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

e<-forPlotting


    options(rgl.useNULL=TRUE)
   plot3d(x=e$Dim.1,y=e$Dim.2,z=e$Dim.3,xlab="", ylab="", zlab="")
        rglwidget()

splt <- split(seq_along(sampleNames), ceiling(seq_along(sampleNames)/20))
for(i in splt){

text3d(x=e$Dim.1[i],y=e$Dim.2[i],z=e$Dim.3[i],text=sampleNames[i])
  }
    rglwidget()
    
    
    
    

```




```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library("Rtsne")


set.seed(42)
d<-Rtsne(proteinMatrixTibble,dims=3)
colnames(d$Y)<-c("Dim.1","Dim.2","Dim.3")
e<-as.data.frame(d$Y)
options(rgl.useNULL=TRUE)
   plot3d(x=e$Dim.1,y=e$Dim.2,z=e$Dim.3,xlab="", ylab="", zlab="",col=cols2)
        rglwidget()

splt <- split(seq_along(sampleNames), ceiling(seq_along(sampleNames)/20))
for(i in splt){

text3d(x=e$Dim.1[i],y=e$Dim.2[i],z=e$Dim.3[i],text=sampleNames[i],col=cols2[i])
  }
    rglwidget()
 

```



while (!is.null(dev.list()))  dev.off()




```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
colnames(e)<-c("aa","bb","cc")
p<-ggplot(data=e)+
    geom_text(aes(x=aa,y=bb,label=sampleNames,color=cols))
ggplotly(p)

```















```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

sampleNames <- row.names(proteinMatrix)

zeroHandle <-sapply(as_tibble(proteinMatrix),function(x){
          unlist(lapply(x,function(y){
              if (y<0.1){     
              
              y<- .001   
              } else{
                  y
              }
          }
          ))
      }
  )



proteinMatrixTibble<-as_tibble(zeroHandle)

proteinMatrixTibble <- proteinMatrixTibble %>% mutate_all(log10)

pcaResults<- PCA(proteinMatrixTibble,graph=FALSE,scale.unit = F)


set.seed(42)
#d<-Rtsne(pcaResults$ind$coord,pca=F,dims=3,perplexity = 50,max_iter=5000)

d<-Rtsne(proteinMatrixTibble,dims=3,perplexity=200,max_iter=2000)


colnames(d$Y)<-c("Axis1","Axis2","Axis3")
e<-as.data.frame(d$Y)

p <- ggplot(data=e)+
    geom_point(aes(x=Axis1,y=Axis2,col=cols))+
    theme_void()+
  ggtitle("t-SNE Plot of Protein Data")


ggplotly(p)

```


There are some weird taxonomic assignments in the t-SNE:

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

tibe<-as_tibble(bind_cols(e,samples=sampleNames,color=cols))
leftx  <- -40
rightx <- -30
boty   <- -10
topy   <- 0


zoomData <- tibe %>% filter(Axis1 > leftx) %>% filter(Axis1 < rightx) %>% filter(Axis2 > boty) %>% filter(Axis2 < topy)


p <- ggplot(data=zoomData)+
    geom_label(aes(x=Axis1,y=Axis2,label=samples),col=zoomData$color)

ggplotly(p)

```




```{r}
options(rgl.useNULL=TRUE)
   plot3d(x=e$Axis1,y=e$Axis2,z=e$Axis3,xlab="", ylab="", zlab="",col=cols2)
        rglwidget()

splt <- split(seq_along(sampleNames), ceiling(seq_along(sampleNames)/20))
for(i in splt){

text3d(x=e$Axis1,y=e$Axis2,z=e$Axis3,text=sampleNames[i],col=cols2[i])
  }
    rglwidget()
```










I created the color column as named colors (eg "Black","Blue"), but plotly needs these to be in HEX


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# cols is a vector of named colors
# First we convert the named colors to rgb, and store the values in a list
rgbVal<-lapply(cols2,col2rgb)
# Then convert the list to a data frame and flip it
rgbVal<-as.data.frame(rgbVal,col.names = seq(1,length(rgbVal),by=1))
# Then we convert the rgb values to HEX
hexVal<-sapply(1:length(rgbVal[,1]),function(x){ rgb(rgbVal[x,"red"],rgbVal[x,"green"],rgbVal[x,"blue"],maxColorValue = 255) })
```





```{r}
plot_ly(x=e$Axis1,y=e$Axis2,z=e$Axis3,text=sampleNames,type="scatter3d",colors=hexVal)
```




















```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

leftx  <- -40
rightx <- -30
boty   <- -4.5
topy   <- 0


zoomData <- tibe %>% filter(Axis1 > leftx) %>% filter(Axis1 < rightx) %>% filter(Axis2 > boty) %>% filter(Axis2 < topy)

p <- ggplot(data=zoomData)+
    geom_label(aes(x=Axis1,y=Axis2,label=samples),col=zoomData$color)

ggplotly(p)

```


# Dendrogram 


Let's look at the dendrogram results to see if this might be real and not just a t-SNE effect...


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
dend<-read_rds("Data/Distance-euclidean_Clustering-ward.D2_UsedPresenceAbsence_SNR-4_PercentPresence-60_LowCut-3000_HighCut-15000.rds")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

dendhc <- as_hclust_fixed(dend)

dendroNetwork(dendhc,treeOrientation = "horizontal",zoom = TRUE,height=8000)




```



[](data/img/Shewanella-Microbacterium.png)




r<-plot_ly(e, x = ~aa, y = ~bb, z = ~cc, type = 'scatter3d', mode = 'markers',text=sampleNames,color = cols)
chart_link = api_create(r, filename="basic3")
chart_link




