---
title: "Dendrogram and Dots: Figuring it Out"
author: "Chase Clark"
date: "January 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load packages:
```{r}
library("tidyverse")
library("dendextend")
library("MALDIquant")

```


# Spectra Evaluation


Get Protein spectra
```{r}
files<-list.files("Data/Peak_Lists",full.names = TRUE)


proteinSpectraFiles<-files[grep("SummedProteinSpectra",files)]

```


load protein spectra
```{r}
proteinSpectra<-sapply(proteinSpectraFiles,readRDS)
```


check for low-intensity spectra
```{r}


maxIntensities<-sapply(proteinSpectra,function(x)max(intensity(x)))


maxIntSorted<-sort(as.vector(maxIntensities))


```



```{r}
a<-as_tibble(maxIntSorted)
w<-as.vector(unlist(a))

```

 
 
Interquartile Range
```{r}
IQR(w)

```

Standard Deviation

```{r}
sd(w)
mad(w)

```


boxplot

```{r}
boxplot(w)

```


plot ordered intensities

```{r}
plot(w)

```






order spectra by intensity, low to high


```{r}

orderedVector<-order(as.vector(maxIntensities))
orderedSpectra<-proteinSpectra[orderedVector]


```

```{r}
par(mfrow=c(4,4))
sapply(orderedSpectra[1:16],plot)

```






Spectra aren't the best, but they are also not so bad.


Let's move on to peaks/ low peak #



```{r}
remove(proteinSpectra)
```



# Peak Evaluation



```{r}
files<-list.files("Data/Peak_Lists",full.names = TRUE)


proteinPeaksFiles<-files[grep("ProteinPeaks",files)]

```


load protein Peaks
```{r}
proteinPeaks<-sapply(proteinPeaksFiles,readRDS)
```


check for low-intensity Peaks
```{r}


maxIntensities<-sapply(unlist(proteinPeaks),function(x)sd(intensity(x)))


maxIntSorted<-sort(as.vector(maxIntensities))


```



```{r}
a<-as_tibble(maxIntSorted)
w<-as.vector(unlist(a))

```

 
 
Interquartile Range
```{r}
IQR(w)

```

Standard Deviation

```{r}
sd(w)
mad(w)

```


boxplot

```{r}
boxplot(w)

```


plot ordered intensities

```{r}
plot(w)

```






order Peaks by intensity, low to high


```{r}

orderedVector<-order(as.vector(maxIntensities))
orderedPeaks<-unlist(proteinPeaks)[orderedVector]


```

```{r}
par(mfrow=c(4,4))
sapply(orderedPeaks[3000:3016],function(x)plot(x,xlim=c(2000,20000)))
```






